---
title: "Assignment3_Bio708"
output: html_document
date: "2026-01-20"
---
See my README (QMEE/README.md) for more info


Install packages
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr) 

```

Bring in my csv file and data frame from assignment 2
- R renames some titles, I know that I could edit this (and remove the empty rows) myself in a spreadsheet, but I am challenging myself in this class to take my data from the software and do all modifications in R, rather than making any edits in excel 
```{r}
Startle_per_20s.csv <- read_csv("Startle_Trial_Stat_per_20s.csv")
max_velocity_per_20s_clean <- readRDS("max_velocity_per_20s.rds")

```

Start with a simple ggplot
- ordered to have control (100% SETBP1 function) on the left and knockout (0% SETBP1 function) on the right
- Grid lines looked cluttered, so opted for medians written on the plot
- Also tried having dots plotted lightly under the boxplot, but it was very distracting. Behaviour data is known to have a lot of noise, so as I increase my sample size it will only get more distracting - leading me to think that a clean boxplot is going to be more useful to me for displaying my data than a plot with every point
- Anchored at 0 (logical)
- Not very information dense (categorical variable on x-axis), but is very clean and clear

- \n to line break
```{r}
ggplot(max_velocity_per_20s_clean,
     aes(x = factor(selection_result,
          levels = c("LD WTSK", "LD WT", "LD KO", "LD KOKO"),
          labels = c("Wildtype \n (SickKids)", "Wildtype \n (In house)", "Hemizygous\nSETBP1 knockout", "Homozygous\nSETBP1 knockout")), 
         y = max_velocity) ) +
    theme_classic()+
     geom_boxplot() +
  labs(x="Genotype", y="Max Velocity (mm/s)")+
coord_cartesian(ylim = c(0, 200)) +
     scale_y_continuous(breaks = seq(0, 200, by = 10))+
                               
  stat_summary(
    fun = median,
    geom = "text",
    fontface = "bold",
    aes(label = round(after_stat(y), 2)),
    vjust = -0.5
    
  )
```


Now a violin plot
- Have a small dataset, so want a graph with points on it
- Want to see if SETBP1 mutants have a wider range of responses than wildtype control fish do (hyper- and hypo- sensitivity to stimulus), so violin shape will be informative
- Graph anchored at 0 (logical)
- Visually, can easily assess that the small sample size of "Wildtype (in house)" is likely driving its shape on the plot (rather than biological difference between groups)
- Anchored at 0 (logical)
- Not very information dense (categorical variable on x-axis)


```{r}
ggplot(max_velocity_per_20s_clean,
       aes(x = selection_result, y = max_velocity)
) +
  geom_violin(trim = FALSE, alpha = 0.4) +
  geom_jitter(width = 0.2, height = 0, alpha = 0.9, size = 1.5) +
  coord_cartesian(ylim = c(0, 250)) +
  scale_x_discrete(
    limits = c("LD WTSK", "LD WT", "LD KO", "LD KOKO"),
    labels = c(
      "Wildtype \n(SickKids)",
      "Wildtype \n(In house)",
      "Hemizygous \nSETBP1 knockout",
      "Homozygous \nSETBP1 knockout"
    ))+
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1))+
  theme_classic()+
  labs(x = "Genotype", y = "Max velocity (mm/s)") 
```


Now graphing the distance traveled metrics (for the next three graphs)
- Hypothesized that wildtype fish will decrease their distance traveled over time following a startle, and SETBP1 mutant fish will not, this is the trend (or absence of) that I am aiming to capture
- Time as the x-axis (instead of my categorical variables), so more information dense
- Achored against 0 (logical)
- Colour coding based on wildtype groups vs control groups. 
     - Blue red palatte is distinguishable for colourblind individuals.  
     - while hue is difficult to distinguist (Cleveland hierarchy) I do not need the reader to distinguish between the two different controls and two different mutants, I only want them to look at differences between controls and mutants, which this accentuates
- Data is plotted on a common axis, which is optimal in the cleveland hierarchy
- Mutants and controls are distinguished not only by colour but also by shape (colour is low on cleveland hierarchy)

```{r}
Startle_per_20s.csv$`Distance moved 2` <-
  as.numeric(Startle_per_20s.csv$`Distance moved 2`)

Startle_per_20s.csv$`Selection Result` <-
  factor(Startle_per_20s.csv$`Selection Result`)


Startle_summary <- Startle_per_20s.csv %>%
  filter(!is.na(`Selection Result`)) %>%
  group_by(`Time bin`, `Selection Result`) %>%
  summarise(
    mean_distance = mean(`Distance moved 2`, na.rm = TRUE),
    .groups = "drop"
  )


Startle_summary$`Selection Result` <- factor(
  Startle_summary$`Selection Result`,
  levels = c("LD WTSK", "LD WT", "LD KO", "LD KOKO"),
  labels = c(
      "Wildtype (SickKids)",
      "Wildtype (In house)",
      "Hemizygous SETBP1 knockout",
      "Homozygous SETBP1 knockout"))
setbp1_palette <- c(
  "Wildtype (SickKids)"        = "#1f78b4",  
  "Wildtype (In house)"        = "#4fa3d1",  
  "Hemizygous SETBP1 knockout" = "#e31a1c",  
  "Homozygous SETBP1 knockout" = "#fb6a4a"   
)
setbp1_shapes <- c(
  "Wildtype (SickKids)"        = 16, 
  "Wildtype (In house)"        = 16,  
  "Hemizygous SETBP1 knockout" = 17, 
  "Homozygous SETBP1 knockout" = 17   
)
ggplot(Startle_summary,
       aes(x = `Time bin`,
           y = mean_distance,
           colour = `Selection Result`,
           group  = `Selection Result`,
           shape = `Selection Result`)
) +
  
  geom_line(linewidth = 1) +
  geom_point(size = 2)+
 coord_cartesian(ylim=c(0,120))+
  scale_x_discrete(
    labels = function(x) {
      starts <- seq(0, 180, by = 20)
      ends   <- seq(20, 200, by = 20)
      paste0(starts, "–", ends)
    }
  ) +
  scale_colour_manual(values=setbp1_palette)+
  scale_shape_manual(values=setbp1_shapes)+
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1))+
labs(x= "Time after startle (20s intervals)", y="Mean distance moved (mm/s)", colour = "Genotype", shape="Genotype")

```

Graphing distance traveled over time again (For analysis, not a visually optimal graph)
- y-axis modified to get a closer look at the difference between groups (y-axis is not anchored against anything logical, but easier to look at angle between points)
- Added confidence intervals
- It is difficult to differentiate each of the different confidence intervals because they overlap so much

```{r}

Startle_summary <- Startle_per_20s.csv %>%
  filter(!is.na(`Selection Result`)) %>%
  group_by(`Time bin`, `Selection Result`) %>%
  summarise(
    mean_distance = mean(`Distance moved 2`, na.rm = TRUE),
    sd_distance   = sd(`Distance moved 2`, na.rm = TRUE),
    n             = n(),
    se            = sd_distance / sqrt(n),
    ci_lower      = mean_distance - 2 * se,
    ci_upper      = mean_distance + 2 * se,
    .groups = "drop"
  )


Startle_summary$`Selection Result` <- factor(
  Startle_summary$`Selection Result`,
  levels = c("LD WTSK", "LD WT", "LD KO", "LD KOKO"),
  labels = c(
      "Wildtype (SickKids)",
      "Wildtype (In house)",
      "Hemizygous SETBP1 knockout",
      "Homozygous SETBP1 knockout"))
ggplot(Startle_summary,
       aes(x = `Time bin`,
           y = mean_distance,
           colour = `Selection Result`,
           fill = `Selection Result`,
           group  = `Selection Result`,
           shape = `Selection Result`)
) +
  geom_ribbon(
    aes(ymin = ci_lower, ymax = ci_upper),
    alpha = 0.2,
    colour = NA
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2)+
 coord_cartesian(ylim=c(60,120))+
  scale_x_discrete(
    labels = function(x) {
      starts <- seq(0, 180, by = 20)
      ends   <- seq(20, 200, by = 20)
      paste0(starts, "–", ends)}
  ) +
   scale_colour_manual(values=setbp1_palette)+
  scale_fill_manual(values = setbp1_palette)+
  scale_shape_manual(values = setbp1_shapes)+
   guides(fill = "none") +
   theme(
    axis.text.x = element_text(angle = 45, hjust = 1))+
labs(x= "Time after startle (20s intervals)", y="Mean distance moved (mm/s)", colour = "Genotype", shape = "Genotype")
```

Final graph of distance over time
- Clearly addresses my hypothesis, because they are on different plots the angle of lines is easy to distinguish 
- Able to distinguish the confidence intervals!
- Not as high on the Cleveland hierarchy because the information is not on a common axis, however if including confidence intervals being able to interpret them is important 
- Y-axis is logical (starts at 0), but have to look closely to see differences in angle 
    - More important to have a logical axis than to try to skew the reader into seeing a change over time
    - In a publication would make graph large so that even on a linear scale anchored at 0 a difference in angle could still be observed

```{r}

ggplot(Startle_summary,
       aes(x = `Time bin`,
           y = mean_distance, colour = `Selection Result`,
           group  = 1)   
) +
 geom_ribbon(
    aes(ymin = ci_lower, ymax = ci_upper, fill = `Selection Result`),
    alpha = 0.2,
    colour = NA
  
  ) +
  geom_line(
    aes(y = mean_distance, colour = `Selection Result`),  
    linewidth = 1
  )+
  facet_wrap(~`Selection Result`) +   
  coord_cartesian(ylim = c(0, 120)) +  
  scale_x_discrete(
    labels = function(x) {
      starts <- seq(0, 180, by = 20)
      ends   <- seq(20, 200, by = 20)
      paste0(starts, "–", ends)}
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")  
  ) +
  labs(
    x = "Time after startle (20s intervals)",
    y = "Mean distance moved (mm/s)"
  )
 
```

