---
title: "Assignment2_File1_Bio708"
output: html_document
date: "2026-01-14"
---

Load important packages
```{r}

library(tidyverse)
library(dplyr)

```

Reading my csv file with my data
- has warning messages because some columns have the same names as eachother
- R renamed them for me so can proceed with some/rename some to be more specific for my use
```{r}
Startle_per_20s <- read_csv("Startle_Trial_Stat_per_20s.csv")

```
Renaming some important columns to be specific for my use
```{r}

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Velocity...6"
   ] <- "max_velocity"

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Distance moved 2"
] <- "dist_moved"


colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Time bin"
     ] <- "time_bin"
```

Checking the R class that columns got added as, then correcting when needed

Check what class selection result variable is
- after thinking, theres only 4 options so may be best as a factor 
```{r}
class(Startle_per_20s$selection_result)

```

convert class selection to factors 
```{r}
Startle_per_20s$selection_result <-
  as.factor(Startle_per_20s$selection_result)
class(Startle_per_20s$selection_result)
```

Check what class velocity is
```{r}
class(Startle_per_20s$max_velocity)

```
Changing velocity to numeric and checking that it worked
- introduces NAs, this is okay because these are spreadsheet lines at the top that are before the dataset has started
- count number of NAs, to make sure I only have the three expected blank lines, and none embedded in the dataset
```{r}
Startle_per_20s$max_velocity <-
    as.numeric(Startle_per_20s$max_velocity)
class(Startle_per_20s$max_velocity)



sum(is.na(Startle_per_20s$max_velocity))
```
Check what class timepoint variable is
- while time is something that could be numeric, in my case I am using it more as a category than a numeric piece of data, so leaving it as a charchter is okay
```{r}
class(Startle_per_20s$time_bin)
```
Check what class distance moved variable is
```{r}
class(Startle_per_20s$dist_moved)

```

Correct class of distance moved variable from charachter to numeric
- This introduces NAs to the dataset
- This is okay because these are empty lines at the top of the sheet that had no data in them
- Check number of NAs to make sure it is only the three that I would expect, and none embedded into the dataset
```{r}
Startle_per_20s$dist_moved <- 
  as.numeric(Startle_per_20s$dist_moved)

class(Startle_per_20s$dist_moved)

sum(is.na(Startle_per_20s$dist_moved))
```
So then getting rid of the three rows where the max velocity column is NA (these are the same as the three rows where distance moved is NA so it removes those as well)

```{r}
Startle_per_20s <- Startle_per_20s %>% filter(!is.na(max_velocity))
```

I expect my distance moved and maximum velocity to always be 0 or greater (no negative numbers) so check that this is correct in my dataset by plotting it 
- binwidth determines the size of the bars, higher number is wider bars so more of them kindof blend together, smaller width can distinguish individual lines better
- This shows that all of my maximum velocity values are positive (as expected so this is good!)
- note to self: at this point am looking at max velocity for all timepoints, not just the first time bin (which is the most biologically relevant)
```{r}
ggplot(
     data = subset(Startle_per_20s),
     aes (x = max_velocity)
 ) +
  geom_histogram(binwidth=10
                 )

```
- For next part of the assignment, convert this to long format, and plot histogram for only the first time bin coloured by selection result (biologically relevant histogram, this is just for data checking)

Plot distance moved for all time bins, I expect distance travelled should always be 0 or greater, if not there is an error in the data 
- histogram shows that distance travelled is always a positive number which is good!
- excludes the three NAs, which is what I want
```{r}
ggplot(data=subset(Startle_per_20s), 
       aes(x=dist_moved)
)+
  geom_histogram(binwidth=5)
       
```

Making a dataframe that just looks at the first timepoint for use looking at maximum velocity
- slice_min takes only the smallest of the variable (here time bin), so it only takes the time 15-15:20 which is the first 30s
  - this is confirmed by visually inspecting the dataframe and i can see that the timepoints are all the 20s that I wanted
- with_ties is so that in the case the same fish has two values that are the lowest time bin (which shouldnt happen given my sampling style), it keeps only one of them
- nrow is a check to make sure that i only got one row per fish (there are 65 fish)
```{r}
max_velocity_per_20s <- Startle_per_20s %>%
  group_by(Trial, Arena) %>%
  slice_min(`time_bin`, n = 1, with_ties = FALSE) %>%
 ungroup()


nrow(max_velocity_per_20s)
view(max_velocity_per_20s)
```

Also want to make a dataframe that is a sum of the distances moved for every time bin per fish to get total maximum distance moved
- use view to look at it like a spreadsheet
- only has columns for dist moved and velocity because they are the columns i changed to be numeric (the is.numeric function only uses the numeric ones)
    - that is okay these are the only ones i want, just good to remember
```{r}
total_dist_moved <- Startle_per_20s %>%
  group_by(selection_result,Trial,Arena) %>%
  summarise(across(where(is.numeric),~sum(.x,na.rm=TRUE)),
  .groups = "drop"
  )

view(total_dist_moved)

```


Save cleaned variables that are set up how i want to use them (max velocity with just first time bin)(distance travelled as the sum of all time bins)
- dont want these "made' objects committed to my repo, so used gitignore (click on it in files, and it opens a new little tab you can write in) to put in *.rmd which means that all rmd files are gitignored
- you can do it file by file, but i thought this would be best so that i dont forget to gitignore any of them 
```{r}
saveRDS(max_velocity_per_20s,"max_velocity_per_20s.rds")
saveRDS(total_dist_moved,"total_dist_moved.rds")

```
