Assignment6_Bio708

Instructions:
Formulate a clear hypothesis about your data.
Make a linear model for your hypothesis.
Draw and discuss at least one of each of the following:
diagnostic plot
prediction plot
inferential plot (e.g., a coefficient plot, or something from emmeans or effects)

Hypothesis (Prediction):
My hypothesis is that both genotypes of SETBP1 knockout zebrafish will display a decreased total distance moved relative to wildtype zebrafish in a distance travelled assay

Linear model:
Chose a one way ANOVA because I have a categorical predictor variable (genotype)nd a continuous outcome variable, and I am trying to see variance in distance



Get set up
```{r}
library(tidyverse)
install.packages("emmeans")
library(emmeans)
install.packages("dotwhisker")
library(dotwhisker)
install.packages("effects")
library(effects)
total_dist_clean <- readRDS("total_dist_moved.rds")
```

Run a one-way ANOVA
```{r}
dist_model <- lm(dist_moved ~ selection_result, data=total_dist_clean)

anova(dist_model)
```

Diagnostic plots:

Residuals vs fitted plot
- no fan shape, this is good because that would show heteroscedasticity
- pretty minimal curve around the line of predicted
- (In vertical lines because categorical, 4 lines for 4 categories)
- (residuals go very high and low, this is because distance data is around 1000)

Q-Q residuals plot
- Is a straight diagonal line, fits reference line well
- A little bit tailed on each end

Scale-location plot:
- line trend pretty horizontal 
- again no fan shape

Overall, these three diagnostic plots suggest a normal distribution, with minimal deviation from model assumptions


Residuals vs leverage plot:
- looks like it has a few points outside range of Cook's distance, so investigating further
- Threshold of 4/n
- 4 points outside this, so removed them and ran a new anova to see if it would make any big changes
- F increased a little and p decreased a little but not a lot
- so this can be attributed to biological variation and not these 4 fish driving the model 

```{r}

plot(dist_model, id.n=4)



cooks <- cooks.distance(dist_model)
summary(cooks)
threshold <- 4 / nrow(total_dist_clean)
which(cooks > threshold)
influential <- which(cooks > 4/nrow(total_dist_clean))
model2 <- lm(dist_moved ~ selection_result,
             data = total_dist_clean[-influential, ])

anova(dist_model)
anova(model2)

```

Prediction plot:

Emmeans:
- blue lines represent confidence intervals
- confidence intervals overlap a lot 
- red lines are pairwise comparisons (mostly put this in to have for future code, I don't need it to do pairwise comparison here)

```{r}
plot(emmeans(dist_model, ~ selection_result), comparisons = TRUE)
```

Effects plot
- this is same as the plot above functionally because I don't have any covariates here (I just want this code in my notes because I will in the future include covariates for batch and sex and body size)
- vertical error bars are 95% CIs for the predicted mean
- CIs overlap substantially

```{r}
plot(allEffects(dist_model))
```
Inferential plot:

Dot-whisker plot
- Right now I don't have data for the main covariates that I want to include here, which would be sex, size, and batch, so am working with arena only
- When have others that will reasonably interact with one another, will swap out for them (I understand that the biological relevance of arena and genotype interactions are minimal, however sex and length are very importnt here so want to make sure I know how to run the code for the future)
- Looking to see if whiskers cross 0, those that do not are statistically significant at the chosen 95% confidence interval
- most of my intervals cross 0, meaning that the sign of the effect is unclear
- Overall, this is not a biologicaly meaningful metric so I won't draw any conclusions from what I see, I just wanted to include something else so I could plot something other than just distance moved by genotype again
```{r}
add_model <- lm(dist_moved ~ selection_result+Arena, data=total_dist_clean)
int_model <- lm(dist_moved ~ selection_result*Arena, data=total_dist_clean)


dotwhisker::dwplot(
              list(additive = add_model, interactive = int_model
                   )) + geom_vline(xintercept=0, lty =2)

```


