---
title: "Untitled"
output: html_document
date: "2026-02-28"
---

Hypothesis (Prediction):
My hypothesis is that both genotypes of SETBP1 knockout zebrafish will display a decreased total distance moved relative to both genotypes of wildtype zebrafish in a distance travelled assay

Overview:
Overall, I think that a LMM is the most appropriate model for my dataset, because I have very important covariates such as batch, sex, and length that need to be accounted for.For the sake of this homework assignment, I have completed a GLM instead.
I have a categorical predictor variable (genotype) and a continuous outcome variable, and I am trying to see variance in distance.
- Because poisson and quasipoisson are for count variables, these are not appropriate here (even using an offset wouldn't be appropriate for this data type)


```{r}
library(tidyverse)
library(DHARMa)
library(see)
library(emmeans)
total_dist_clean <- readRDS("total_dist_moved.rds")
```


Model - GLM with a gaussian family and an identity link
- this in essence is the same as the ANOVA I ran last week 
- used a gaussian distribution because I am starting off with the assumption that the results will follow a normal distribution
- used an identity link, this means that the expected value of Y is equal to the linear predictor itself. This is me assuming that the errors are normally distributed.
- there is no transformation applied in this model

- residual deviance/degrees of freedom is very high, but that rule for overdispersion only applies for poisson/binomial families so don't need to worry about it
```{r}
dist_glm_gaus_iden <- glm(dist_moved~selection_result, data=total_dist_clean, family=gaussian(link ="identity") )
summary(dist_glm_gaus_iden)

```

Diagnostic plots:

Residuals vs fitted
- line follows predicted very closely

Q-Q residuals
- a straight diagonal line, fits reference line well
- a little bit tailed at the top, but is minimal, overall still fits well

Scale-location
- line pretty horizontal, showing consistent residual variance
- a bit harder to visualize because data is set in lines by the 4 genotypes, but no crazy concerning trend jumping out on the right


Residuals vs leverage
- inside Cook's distance
  
Posterior predictive check
-the model predicted lines peak a little bit earlier and lower than the observed line, it still captures the general shape well but shows some model imperfection (the shape is still very close and the tails are well matched )

Linearity
- reference line is flat and horizontal (I love it when they tel you right on the plot what you want it to look like)

Homogeneity of variance
- not a straight horizontal line, but looking at the actual numbers of the range on the y-axis (0.6-1.1) this isn't a huge change, just mild heteroscedasticity
- might be seeing this as a result of having 4 categories for predictors and a continuous outcome, it is putting my points into only the 4 bands

Influential observations
- all points are within the contour lines

Normality of residuals
- all points except for one are within the range of variation that you would expect for normal data (grey area). 
- doesn't follow center line super well, it starts high - this may show slight skewedness, wherethe most low residual is less extremem than predicted
- I have a small sample size (total n=65) so some skew isn't crazy to see, esp at the extreme ends that are the most unstable
-an imperfect fit here isn't enough to suggest that I should switch to another kind of GLM (ofc, my data really calls for a LMM)


```{r}
plot(dist_glm_gaus_iden)
performance::check_model(dist_glm_gaus_iden)
```

Inferential plot:

Estimated marginal means (emm) plot
- group means and 95% CI
- all 4 groups overlap a lot, this shows that the difference in distance traveled between groups is statistically unclear
```{r}
emm <- emmeans(dist_glm_gaus_iden, ~ selection_result)
plot(emm)
```

JD: Reasonable, but this is not a “generalized” model in any sense. It's exactly the same as a linear model. Better to use the emmeans comparison arrows (or make contrasts) if you are trying to make an inferential plot. Grade 1.8/3

