---
title: "Assignment2_File2_Bio708"
output: html_document
date: "2026-01-16"
---

Install important packages and load in my RDS files
- class function just to make sure they are dataframes (and i didnt accidentally load in just old variables or anything weird)
```{r}
library(tidyverse)

total_dist_clean <- readRDS("total_dist_moved.rds")
max_velocity_per_20s_clean <- readRDS("max_velocity_per_20s.rds")

class(total_dist_clean)
class(max_velocity_per_20s_clean)

```


Calculate mean dist travelled by genotype
- also calculate the standard deviation (sd) and standard error (se)
```{r}



mean_dist_by_genotype <- total_dist_clean %>%
  group_by(selection_result)%>%
  summarise(
         mean_total_dist = mean(dist_moved),
        sd_total_dist = sd(dist_moved),  
        se_total_dist = sd(dist_moved)/sqrt(n()),
      .groups = "drop" )

print(mean_dist_by_genotype)
```

Calculate mean maximum velocity per genotype
- also calculate standard deviation and standard error

```{r}
max_velocity_per_20s_by_genotype <- max_velocity_per_20s_clean %>%
  group_by(selection_result)%>%
  summarise(
         mean_max_velocity_per_20s = mean(max_velocity),
        sd_total_dist = sd(max_velocity),  
        se_total_dist = sd(max_velocity)/sqrt(n()),
      .groups = "drop" )

print(max_velocity_per_20s_by_genotype)
```

<!-- mark: 2 -->
