---
title: "Assignment2_Bio708"
output: html_document
date: "2026-01-14"
---


Load important packages
```{r}

library(tidyverse)
library(dplyr)

```

Reading my csv file with my data
- has warning messages because some columns have the same names as eachother
- R renamed them for me so can proceed with some/rename some to be more specific for my use
```{r}
Startle_per_20s <- read_csv("Startle_Trial_Stat_per_20s.csv")

```
Renaming some important columns to be specific for my use
```{r}

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Velocity...6"
   ] <- "max_velocity"

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Distance moved 2"
] <- "dist_moved"


colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Time bin"
     ] <- "time_bin"
```


Checking the R class that columns got added as, then correcting when needed

Check what class selection result variable is
- want this as charachter
```{r}
class(Startle_per_20s$selection_result)
```

Check what class velocity is
```{r}
class(Startle_per_20s$max_velocity)

```
Changing velocity to numeric and checking that it worked
- introduces NAs, this is okay because these are spreadsheet lines at the top that are before the dataset has started
- count number of NAs, to make sure I only have the three expected blank lines, and none embedded in the dataset
```{r}
Startle_per_20s$max_velocity <-
    as.numeric(Startle_per_20s$max_velocity)
class(Startle_per_20s$max_velocity)


sum(is.na(Startle_per_20s$max_velocity))
```



Check what class timepoint variable is
- while time is something that could be numeric, in my case I am using it more as a category than a numeric piece of data, so leaving it as a charchter is okay
```{r}
class(Startle_per_20s$time_bin)
```
Check what class distance moved variable is
```{r}
class(Startle_per_20s$dist_moved)

```

Correct class of distance moved variable from charachter to numeric
- This introduces NAs to the dataset
- This is okay because these are empty lines at the top of the sheet that had no data in them
- Check number of NAs to make sure it is only the three that I would expect, and none embedded into the dataset
```{r}
Startle_per_20s$dist_moved <- 
  as.numeric(Startle_per_20s$dist_moved)

class(Startle_per_20s$dist_moved)

sum(is.na(Startle_per_20s$dist_moved))
```


I expect my distance moved and maximum velocity to always be 0 or greater (no negative numbers) so check that this is correct in my dataset by plotting it 
- Has a warning message that it removes my three NAs which is okay
- binwidth determines the size of the bars, higher number is wider bars so more of them kindof blend together, smaller width can distinguish individual lines better
- This shows that all of my maximum velocity values are positive (as expected so this is good!)
- note to self: at this point am looking at max velocity for all timepoints, not just the first time bin (which is the most biologically relevant)
```{r}
ggplot(
     data = subset(Startle_per_20s),
     aes (y = max_velocity)
 ) +
  geom_histogram(binwidth=10
                 )
```
- For next part of the assignment, convert this to long format, and plot histogram for only the first time bin coloured by selection result (biologically relevant histogram, this is just for data checking)

Plot distance moved for all time bins, I expect distance travelled should always be 0 or greater, if not there is an error in the data 
- histogram shows that distance travelled is always a positive number which is good!
- excludes the three NAs, which is what I want
```{r}
ggplot(data=subset(Startle_per_20s), 
       aes(y=dist_moved)
)+
  geom_histogram(binwidth=10)
       
```

