---
title: "Untitled"
output: html_document
date: "2026-02-28"
---

Hypothesis (Prediction):
My hypothesis is that both genotypes of SETBP1 knockout zebrafish will display a decreased total distance moved relative to both genotypes of wildtype zebrafish in a distance travelled assay

Overview:
Overall, I think that a LMM is the most appropriate model for my dataset, because I have very important covariates such as batch, sex, and length that need to be accounted for.For the sake of this homework assignment, I have completed a GLM instead.
I have a categorical predictor variable (genotype) and a continuous outcome variable, and I am trying to see variance in distance.
- Because poisson and quasipoisson are for count variables, these are not appropriate here (even using an offset wouldn't be appropriate for this data type)


```{r}
library(tidyverse)
library(DHARMa)
library(see)
library(emmeans)
total_dist_clean <- readRDS("total_dist_moved.rds")
```


Model - GLM with a gaussian family and an identity link
- this in essence is the same as the ANOVA I ran last week 
- used a gaussian distribution because I am starting off with the assumption that the results will follow a normal distribution
- used an identity link, this means that the expected value of Y is equal to the linear predictor itself. This is me assuming that the errors are normally distributed.
- there is no transformation applied in this model

- residual deviance/degrees of freedom is very high, but that rule for overdispersion only applies for poisson/binomial families so don't need to worry about it
```{r}
dist_glm_gaus_iden <- glm(dist_moved~selection_result, data=total_dist_clean, family=gaussian(link ="identity") )
summary(dist_glm_gaus_iden)

```

Diagnostic plots:
```{r}
plot(dist_glm_gaus_iden)
performance::check_model(dist_glm_gaus_iden)
```

Inferential plots:

Estimated marginal means (emm) plot
- group means and 95% CI
- all 4 groups overlap a lot, this shows that the difference in distance travelled between groups is statistically unclear
```{r}
emm <- emmeans(dist_glm_gaus_iden, ~ selection_result)
plot(emm)
```







