---
title: "Assignment2_Bio708"
output: html_document
date: "2026-01-14"
---


Load important packages
```{r}

library(tidyverse)
library(dplyr)

```

Reading my csv file with my data
- has warning messages because some columns have the same names as eachother
- R renamed them for me so can proceed with some/rename some to be more specific for my use
```{r}
Startle_per_20s <- read_csv("Startle_Trial_Stat_per_20s.csv")

```
Renaming some important columns to be specific for my use
```{r}

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Velocity...6"
   ] <- "max_velocity"

 colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Distance moved 2"
] <- "dist_moved"


colnames(Startle_per_20s)[
     colnames(Startle_per_20s) == "Time bin"
     ] <- "time_bin"
```


Checking the R class that columns got added as, then correcting when needed

Check what class selection result variable is
- want this as charachter
```{r}
class(Startle_per_20s$selection_result)
```

Check what class velocity is
```{r}
class(Startle_per_20s$max_velocity)

```
Changing velocity to numeric and checking that it worked
- introduces NAs, this is okay because these are spreadsheet lines at the top that are before the dataset has started
- Maybe should add in a line that counts the number of NAs added to confirm this??
```{r}
Startle_per_20s$max_velocity <-
    as.numeric(Startle_per_20s$max_velocity)
class(Startle_per_20s$max_velocity)
```



Check what class timepoint variable is
- while time is something that could be numeric, in my case I am using it more as a category than a numeric peice of data, so leaving it as a charchter is okay
```{r}
class(Startle_per_20s$time_bin)
```
Check what class distance moved variable is
```{r}
class(Startle_per_20s$dist_moved)

```

Correct class of distance moved variable from charachter to numeric
- This introduces NAs to the dataset
- This is okay because these are empty lines at the top of the sheet that had no data in them
```{r}
Startle_per_20s$dist_moved <- 
  as.numeric(Startle_per_20s$dist_moved)

class(Startle_per_20s$dist_moved)
```

