I did all my work in the console by accident, so putting it in r markdown now to be able to commit to github

```{r}
## Bio708 Assignment 1: Embryo Startle Response Assay
library(readxl)
Startle_Trial_Stat_per_20s <- read_excel("Startle Trial Stat per 20s.xlsx")
```

Check that it’s a data frame
```{r}
class(Startle_Trial_Stat_per_20s)
```

Attach packages 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

```

Renaming my excel titles so theyre easier to use
```{r}
Startle_Trial_Stat_per_20s <- Startle_Trial_Stat_per_20s
 colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Velocity...6"
   ] <- "max_velocity"

colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Time bin"
     ] <- "time_bin"

```

want to look at max velocity immediately following the startle, so only look at the first time bin
```{r}
first20s <- "0:15:00-0:15:20"
```

also need to convert from categorical to numerical (introduces na from the first two columns which are words, I think this is ok bc they aren’t result data, just specifying the units the rest of the sheet uses)

```{r}
Startle_Trial_Stat_per_20s$max_velocity <-
    as.numeric(Startle_Trial_Stat_per_20s$max_velocity)

```

now boxplot, with the median values on the plot
notes for future:
- alpha is how light or dark the dots are
- ylim is how high axis goes, the scale_y_continuous stuff is how i labelled it the way i want
- stat summary reruns the calculation for median (but is the same value as the median of boxplot because its the same calculation), which is why you have to respecify with the function line that you want the median (could also probably put mean there if wanted )
- rounded to 2 decimal points with the 2 in the aes line
-vjust line is to put the label above the value, so where on the boxplot the written out median is placed
```{r}
ggplot(
     data = subset(Startle_Trial_Stat_per_20s, time_bin == "0:15:00-0:15:20"),
     aes(x = factor(selection_result), y = max_velocity)
 ) +
     geom_boxplot() +
   geom_jitter(alpha = 0.7)+
coord_cartesian(ylim = c(0, 200)) +
     scale_y_continuous(breaks = seq(0, 200, by = 10))+
  stat_summary(
    fun = median,
    geom = "text",
    aes(label = round(after_stat(y), 2)),
    vjust = -0.5)


```




Now going to graph distance traveled
Introduces some NA, these are just blanks on the sheet where other lines had details about units etc. so dont need them

```{r}
 colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Distance moved 2"
] <- "dist_moved"

Startle_Trial_Stat_per_20s$dist_moved <- 
  as.numeric(Startle_Trial_Stat_per_20s$dist_moved)
```

now only want to look at the total distance traveled per individual
am purposely keeping the grouping, because i want it grouped for my graph and for my analysis
If i wanted to ungroup it could put in .groups = "drop"

Also getting rid of first three rows where seection result and distance moved are both NA
```{r}
library(dplyr)

total_dist <-summarise(
  group_by(Startle_Trial_Stat_per_20s ,Trial, Arena, selection_result),
  total_dist_moved = sum(dist_moved)
  
)

total_dist <- total_dist[
  !is.na(total_dist$selection_result),
]

```

now boxplot total distance moved, with the median values printed on the plot
warning is okay, it removes the three blank lines which i want removed
```{r}
ggplot(
  data = total_dist,
  aes(
    x = factor(selection_result),
    y = total_dist_moved
  )
) +
  geom_boxplot() +
  geom_jitter(alpha = 0.7)+

  stat_summary(
    fun = median,
    geom = "text",
    aes(label = round(after_stat(y), 2)),
    vjust = -0.5)


```



