Assignment6_Bio708

Instructions:
Formulate a clear hypothesis about your data.
Make a linear model for your hypothesis.
Draw and discuss at least one of each of the following:
diagnostic plot
prediction plot
inferential plot (e.g., a coefficient plot, or something from emmeans or effects)

Hypothesis (Prediction):
My hypothesis is that both genotypes of SETBP1 knockout zebrafish will display a decreased total distance moved relative to wildtype zebrafish in a distance travelled assay

Linear model:
Chose a one way ANOVA because I have a categorical predictor variable (genotype), and a continuous outcome variable, and I am trying to see variance in distance



Get set up
```{r}
library(tidyverse)
install.packages("emmeans")
library(emmeans)
library(effects)
total_dist_clean <- readRDS("total_dist_moved.rds")
```

Run a one-way ANOVA
```{r}
dist_model <- lm(dist_moved ~ selection_result, data=total_dist_clean)

anova(dist_model)
```

Diagnostic plots:

Residuals vs fitted plot
- no fan shape, this is good because that would show heteroscedasticity
- pretty minimal curve around the line of predicted
- (In vertical lines because categorical, 4 lines for 4 categories)
- (residuals go very high and low, this is because distance data is around 1000)

Q-Q residuals plot
- Is a straight diagonal line, fits reference line well
- A little bit tailed on each end

Scale-location plot:
- line trend pretty horizontal 
- again no fan shape

Overall, these three diagnostic plots suggest a normal distribution, with minimal deviation from model assumptions


Residuals vs leverage plot:
- looks like it has a few points outside range of Cook's distance, so investigating further
- Threshold of 4/n
- 4 points outside this, so removed them and ran a new anova to see if it would make any big changes
- F increased a little and p decreased a little but not a lot
- so this can be attributed to biological variation and not these 4 fish driving the model 

```{r}

plot(dist_model, id.n=4)



cooks <- cooks.distance(dist_model)
summary(cooks)
threshold <- 4 / nrow(total_dist_clean)
which(cooks > threshold)
influential <- which(cooks > 4/nrow(total_dist_clean))
model2 <- lm(dist_moved ~ selection_result,
             data = total_dist_clean[-influential, ])

anova(dist_model)
anova(model2)

```

Prediction plot:
- blue lines represent confidence intervals
- confidence intervals overlap a lot 
- red lines are pairwise comparisons (mostly put this in to have for future code, I don't need it to do pairwise comparison here)

```{r}
plot(emmeans(dist_model, ~ selection_result), comparisons = TRUE)
```

Inferential plots:

Effects plot
```{r}

```

