---
title: "Assignment7_Bio708"
output: html_document
date: "2026-02-28"
---

Instructions:
  Make a generalized linear model for one or more of your hypotheses. Draw and discuss at least one of each of the following:

diagnostic plot
inferential plot (e.g., a coefficient plot, or something from emmeans or effects)

```



```{r}
library(tidyverse)
library(DHARMa)
#install.packages("see", dependencies = TRUE)
library(see)
total_dist_clean <- readRDS("total_dist_moved.rds")
```





```{r}
dist_glm_gaus_iden <- glm(dist_moved~selection_result, data=total_dist_clean, family=gaussian(link ="identity") )
summary(dist_glm_gaus_iden)
plot(dist_glm_gaus_iden)
performance::check_model(dist_glm_gaus_iden)
```




Lets try Gamma and log 
```{r}
dist_glm <- glm(dist_moved~selection_result, data=total_dist_clean, family=Gamma(link ="log") )
summary(dist_glm)
plot(dist_glm)
```

```{r}
performance::check_model(dist_glm)
```



Now lets try gaussian with log link
```{r}
dist_glm_gauss <- glm(dist_moved~selection_result, data=total_dist_clean, family=gaussian(link ="log") )
summary(dist_glm_gauss)
plot(dist_glm_gauss)

```
```{r}
performance::check_model(dist_glm_gauss)
```

now quasipoisson and log link
```{r}
dist_glm_qp <- glm(dist_moved~selection_result, data=total_dist_clean, family=quasipoisson(link ="log") )
summary(dist_glm_qp)
plot(dist_glm_qp)
```

```{r}
performance::check_model(dist_glm_qp)
```


```{r}
dist_glm_gau <- glm(dist_moved~selection_result, data=total_dist_clean, family=gaussian(link ="log") )
summary(dist_glm_gau)
plot(dist_glm_gau)
```

