I did all my work in the console by accident, so putting it in r markdown now to be able to commit to github

```{r}
## Bio708 Assignment 1: Embryo Startle Response Assay
library(readxl)
Startle_Trial_Stat_per_20s <- read_excel("Startle Trial Stat per 20s.xlsx")
```

Check that it’s a data frame
```{r}
class(Startle_Trial_Stat_per_20s)
```

Attach package dplyr and tidyr
```{r}
library(dplyr)
library(tidyr)

```

Renaming my excel titles so theyre easier to use
```{r}
Startle_Trial_Stat_per_20s <- Startle_Trial_Stat_per_20s
 colnames(Startle_Trial_Stat_per_20s)[
+     colnames(Startle_Trial_Stat_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_Trial_Stat_per_20s)[
+     colnames(Startle_Trial_Stat_per_20s) == "Velocity...6"
   ] <- "max_velocity"

colnames(Startle_Trial_Stat_per_20s)[
+     colnames(Startle_Trial_Stat_per_20s) == "Time bin"
     ] <- "time_bin"

```

want to look at max velocity immediately following the startle, so only look at the first time bin
```{r}
first20s <- "0:15:00-0:15:20"
```

also need to convert from categorical to numerical (introduces na from the first two columns which are words, I think this is ok bc they aren’t result data, just specifying the units the rest of the sheet uses)

```{r}
Startle_Trial_Stat_per_20s$max_velocity <-
    as.numeric(Startle_Trial_Stat_per_20s$max_velocity)

```

now boxplot
```{r}

```

