I did all my work in the console by accident, so putting it in r markdown now to be able to commit to github

```{r}
## Bio708 Assignment 1: Embryo Startle Response Assay
library(readxl)
Startle_Trial_Stat_per_20s <- read_excel("Startle Trial Stat per 20s.xlsx")
```

Check that it’s a data frame
```{r}
class(Startle_Trial_Stat_per_20s)
```

Attach package dplyr and tidyr
```{r}
library(dplyr)
library(tidyr)

```

Renaming my excel titles so theyre easier to use
```{r}
Startle_Trial_Stat_per_20s <- Startle_Trial_Stat_per_20s
 colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Selection Result"
] <- "selection_result"

colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Velocity...6"
   ] <- "max_velocity"

colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Time bin"
     ] <- "time_bin"

```

want to look at max velocity immediately following the startle, so only look at the first time bin
```{r}
first20s <- "0:15:00-0:15:20"
```

also need to convert from categorical to numerical (introduces na from the first two columns which are words, I think this is ok bc they aren’t result data, just specifying the units the rest of the sheet uses)

```{r}
Startle_Trial_Stat_per_20s$max_velocity <-
    as.numeric(Startle_Trial_Stat_per_20s$max_velocity)

```

now boxplot
```{r}
ggplot(
     data = subset(Startle_Trial_Stat_per_20s, time_bin == "0:15:00-0:15:20"),
     aes(x = factor(selection_result), y = max_velocity)
 ) +
     geom_boxplot() +
   geom_jitter(alpha = 0.7)+
coord_cartesian(ylim = c(0, 200)) +
     scale_y_continuous(breaks = seq(0, 200, by = 10))


```

Now going to graph distance traveled
Introduces some NA, these are just blanks on the sheet where other lines had details about units etc. so dont need them

```{r}
 colnames(Startle_Trial_Stat_per_20s)[
     colnames(Startle_Trial_Stat_per_20s) == "Distance moved 2"
] <- "dist_moved"

Startle_Trial_Stat_per_20s$dist_moved <- 
  as.numeric(Startle_Trial_Stat_per_20s$dist_moved)
```

now only want to look at the total distance traveled per individual
am purposely keeping the grouping, because i want it grouped for my graph and for my analysis
If i wanted to ungroup it could put in .groups = "drop"

grouping is wrong for now, doesnt sum properly

```{r}
library(dplyr)

total_dist <-summarise(
  group_by(Startle_Trial_Stat_per_20s ,Trial, Arena),
  total_dist_moved=sum(dist_moved)
)



```

now boxplot
warning is okay, it removes the three blank lines which i want removed
```{r}
ggplot(
     data = subset(Startle_Trial_Stat_per_20s),
     aes(x = factor(selection_result), y = total_dist)
 ) +
    geom_boxplot() +
    geom_jitter(alpha = 0.7)


```

